---
title: "Joins and strings and dates"
format: html
editor: visual
---
```{r}
library(tidyverse)
library(palmerpenguins)
library(lubridate)
```

## Joins
```{r}
animals <- tibble::tribble(
             ~location,   ~species,  ~maturity,
              "lagoon",   "bobcat",    "adult",
               "bluff",   "coyote", "juvenile",
               "creek",      "fox",    "adult",
                "oaks", "squirrel", "juvenile",
               "bluff",   "bobcat",    "adult"
             )

sites <- tibble::tribble(
           ~location,      ~full_site_name, ~jurisdiction,
             "beach",       "Goleta Beach",     "SB City",
            "lagoon",        "UCSB Lagoon",        "UCSB",
             "bluff",       "Ellwood Mesa",     "SB City",
              "oaks", "Fremont Campground",        "USFS"
           )
```

### Full join

```{r}
full_join_example <- full_join(animals, sites, by = "location")
```

### Left join

```{r}
left_join_example <- left_join(animals, sites, by = "location")
```

### Inner join

```{r}
inner_join_example <- inner_join(animals, sites, by = "location")
```

```{r}
sites_2 <- sites %>%
  rename(loc = location)
```

Example of joining when key variables have non-matching names

```{r}
nonmatching <- full_join(animals, sites_2, by = c("location" = "loc"))
```

### Semi join

```{r}
semi_join_example <- semi_join(animals, sites, by = "location")
```

### Anti join

```{r}
anti_join_example <- anti_join(animals, sites, by = "location")
```

## Part 1: Data wrangling refreshers

Refresher 1: Starting with the penguins dataset in the palmerpenguins package, write a single piped sequence in which you:

Only include penguins at Biscoe and Dream islands
Remove the year and sex variables
Add a new column called body_mass_kg, with penguin mass converted from grams to kilograms
Rename the island variable to location

```{r}
refresher_1 <- penguins %>%
  filter(island %in% c("Biscoe", "Dream")) %>%
  select(-year, -sex) %>%
  mutate(body_mass_kg = body_mass_g / 1000) %>%
  rename(location = island)
```


Refresher 2: Staring with the penguins dataset in the palmerpenguins package, write a single piped sequence in which you:

Limit to only Adelie penguins
Remove any observations where flipper_length_mm is NA (hint: !is.na())
Group the data by sex
Find the mean (mean()), standard deviation (sd) and sample size (n) of flipper_length_mm for male and female Adelie penguins, returned in a nice summary table



## Dates and times with `lubridate`

```{r}
my_date <- "03-15-1998"

# mdy, ymd, dmy
lubridate::mdy(my_date)

my_date <- "11/7/2003"
lubridate::dmy(my_date)

my_date <- "08-Jun-1974"
lubridate::dmy(my_date)

my_date <- "19610518"
my_real_date <- lubridate::ymd(my_date)

lubridate::ymd("1942-08-30")
lubridate::mdy("09/22/84")

time <- "2020-08-12 11:18"
time <- lubridate::ymd_hm(time)

lubridate::with_tz(time, "America/Los_Angeles")

lubridate::with_tz(time, "Asia/Jakarta")

my_time <- lubridate::ymd_hms(time)
day(my_time)
month(my_time)
hour(my_time)
second(my_time)
minute(my_time)

teddy_bday <- lubridate::ymd("20170615")
wday(teddy_bday, label = TRUE)
week(teddy_bday)

wday(lubridate::ymd("20000322"), label = TRUE)
```

